---
title: "Anleitung zur Statistische Analyse"
author: "Mohan Bi"
date: "2025-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing generating datasets 
#Not always using linear model!!


```{r package}

# Bibliothek laden
library(ggplot2)

```

###Butterfly population and sunspots
#Context
A group of ecologists is studying factors influencing biomass production in a particular grassland ecosystem. They collect data on various factors, including butterfly population, sunspots, soil pH and biomass production over a period of time.
#Findings
After running a linear regression, they find a statistically significant correlation between butterfly population and biomass production. Surprisingly, they also find that the number of sunspots observed in a given year is correlated with biomass production.
#Why is this a spurious correlation?
- Butterfly population fluctuations could be linked to temperature, seasonal changes, or other indirect ecological factors but are not a direct cause of biomass increase.
- Sunspots have no biological link to plant growth, but both sunspots and biomass may independently be influenced by climatic patterns like precipitation and temperature.
- An overlooked common cause (e.g., seasonal climate changes) might be driving both butterfly abundance and plant biomass, creating the illusion of a causal link.

```{r Data_import}
# Butterfly population count (randomized, 100 observations)
butterfly_population <- c(179, 85, 235, 218, 251, 492, 382, 250, 338, 393,
                          225, 160, 106, 87, 348, 275, 243, 204, 80, 135,
                          342, 102, 127, 279, 147, 196, 206, 337, 89, 238,
                          105, 151, 150, 495, 389, 343, 303, 354, 290, 356,
                          138, 329, 317, 457, 241, 332, 383, 308, 420, 187,
                          155, 334, 285, 491, 480, 157, 488, 350, 203, 207,
                          68, 259, 150, 71, 201, 264, 324, 244, 467, 116,
                          385, 263, 73, 91, 200, 318, 467, 222, 289, 478,
                          170, 186, 467, 484, 473, 109, 365, 277, 285, 267,
                          310, 309, 286, 495, 222, 330, 78, 90, 64, 278)

# Sunspots count (randomized, 100 observations)
sunspots <- c(62, 31, 75, 94, 11, 55, 82, 76, 44, 99,
              71, 53, 14, 21, 77, 88, 80, 20, 31, 54,
              40, 38, 59, 69, 83, 96, 12, 90, 36, 79,
              65, 72, 98, 85, 39, 67, 42, 92, 97, 46,
              22, 91, 24, 49, 87, 32, 26, 58, 61, 34,
              33, 84, 66, 28, 70, 17, 37, 30, 23, 47,
              20, 55, 41, 68, 15, 16, 60, 50, 35, 57,
              45, 86, 48, 43, 27, 13, 95, 56, 63, 29,
              74, 25, 19, 18, 81, 89, 93, 10, 9, 7,
              73, 52, 8, 3, 6, 5, 4, 2, 1, 0)

# Biomass (randomly generated with spurious correlation to butterfly_population and sunspots)
set.seed(99)
random_noise <- rnorm(100, mean = 0, sd = 1)
biomass <- 20 + 0.1 * butterfly_population + 0.05 * sunspots + random_noise
```



```{r dataframe}
set.seed(99)
soil_ph <- runif(n = 100, min = 3, max = 10)
random_noise <- rnorm(100, mean = 0, sd = 1)
butterfly_population <- c(179, 85, 235, 218, 251, 492, 382, 250, 338, 393,
                          225, 160, 106, 87, 348, 275, 243, 204, 80, 135,
                          342, 102, 127, 279, 147, 196, 206, 337, 89, 238,
                          105, 151, 150, 495, 389, 343, 303, 354, 290, 356,
                          138, 329, 317, 457, 241, 332, 383, 308, 420, 187,
                          155, 334, 285, 491, 480, 157, 488, 350, 203, 207,
                          68, 259, 150, 71, 201, 264, 324, 244, 467, 116,
                          385, 263, 73, 91, 200, 318, 467, 222, 289, 478,
                          170, 186, 467, 484, 473, 109, 365, 277, 285, 267,
                          310, 309, 286, 495, 222, 330, 78, 90, 64, 278)
biomass = -2*soil_ph *soil_ph  +28*soil_ph +random_noise + 0.1* butterfly_population -50
sunspots <- as.integer(0.1*biomass + 50 + random_noise*15)

df <- data.frame(
  sample_number = c(1:100),
  biomass = biomass,
  butterfly_population = butterfly_population,
  soil_ph = soil_ph,
  sunspots = sunspots,
  stringsAsFactors = FALSE)
```

```{r stat}
summary(biomass)
summary(butterfly_population)
summary(sunspots)
```

#Correlation test
```{r cor}
t.test(df$butterfly_population, df$biomass, method = "pearson")   # Simple t-test
cor.test(df$butterfly_population, df$biomass, method = "spearman")  # Spearman-Korrelation
```

#Perform linear regression
```{r linear regression}
df_s <- df[c(6, 13, 29, 36, 44, 52, 61, 79, 86, 94),]

model <- lm(data = df_s, biomass ~ butterfly_population)
summary(model)

```

#Visualization
```{r visual}
library(ggplot2)

ggplot(df, aes(x=butterfly_population, y=biomass)) +
  geom_point() +
  geom_smooth(method="lm", col="blue") 

ggplot(df, aes(x=soil_ph, y=biomass)) +
  geom_point() +
  geom_smooth(method="lm", col="blue")

p<-ggplot(df, aes(x=butterfly_population, y=biomass)) +
         geom_point() +
         geom_smooth(method="lm", col="blue")+
  xlim(50,500)+
  ylim(20,110)

p_s<-ggplot(df_s, aes(x=butterfly_population, y=biomass)) +
         geom_point() +
         geom_smooth(method="lm", col="red")+
  xlim(50,500)+
  ylim(20,110)
p+p_s+
  plot_layout(ncol = 1)
```

```{r quadratic Model}
quadratic_model <- lm(biomass ~ soil_ph + I(soil_ph^2), data = df)
summary(quadratic_model)

predicted_produ <- predict(
  quadratic_model,
  newdata = data.frame(
    soil_ph = seq(3,
                  10,
                  length.out = 1000
                  )
  )
)

plot(
  df$soil_ph,
  df$biomass,
  xlab = "soil_ph",
  ylab = "biomass"
)
lines(seq(3,
          10,
          length.out = 1000),
      predicted_produ, col = "blue"
      )
```